---
title: "R Beginnings"
output: html_document
---
---
title: "R Beginnings"
output: html_document
---

+++++++++++++++++++++++++++++++++  READ HERE FIRST +++++++++++++++++++++++++++++++++++++++++++++++++++
When you download this document into Rstudio you should see a button that says 'knit' directly above (next to the magnifying glass), if you click the pull down arrow  
there's an option to generate pdf or html. If you generate the html you will get another window pop open with this in a slightly easier read format that  
you can read along while working with the first.R file directly.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Understanding R atomic elements  
R comes with basic data types, these store values such as 1, 1.25, 1e-06, "test", TRUE and other values of this nature.  
the basic types are:  
**Numeric** this one represents numbers, it includes numbers that are both real or integer such as 1, 2, 1.2,1.5  
**Integer** this is numbers that have no decimals only such as 1, 2, 3 they will always end in an L so 1 would be 1L if it's an integer instead of a numeric. If you use an integer it will use less memory, not as important with today's computers but you might run into it in some code.
**Character** these are words, single characters, etc. They will be surrounded by quotes. Examples: "A","b","This is a character string"  
**Logical** this represents true and false conditions. They are represented as TRUE or FALSE (all caps)    
**Complex** complex numbers involving real and imaginary numbers.  

more information available here: https://swcarpentry.github.io/r-novice-inflammation/13-supp-data-structures/ 
To accompany this markdown there is also "first.R" that only has the R code so you can experiment further with that file.  
You will notice that in R the <- sequence is used to assign values, just train your brain to think of this as assigning the values to the right  
to whatever is "pointed" to by the arrow rather than using = signs. For functions you will often use the = sign for passing arguments.  
In the next section of R code I use a function called typeof(), this function will return what the object is, such as  
the basic types mentioned above.  

```{r atomic-types}
x <- 10 # What type is this? Numeric or Integer?
typeof(x)
y <- 10L
typeof(y)
z <- 1+4i
typeof(z)

# Fill out below this line to try the other 2 data types.

```

## Vectors  
An important data type in R is the vector, you can think of it as a single column in excel/open office. You make vectors with the c('value1','value2','etc') statement.  
some examples of vectors are in the next r block, also you should practice making a few others these are used a lot when working in R  
at least try and make one that has only TRUE, FALSE values in it and one with a series of numbers and finally  
one that is a sequence of strings, maybe "bob","sam","bozo" if you are stuck for words.  

```{r vectors}
my_names <- c("Spongebob","Patrick","Plankton","Karen")
my_names
my_ints <- c(1,2,3)
my_ints
```

## Data frames  

The other data type that you will use for most of your work are data frames. You can think of data frames as an excel spreadsheet, they have columns  
and rows and different types of data stored in each column. All columns must have the same data in them but each column can be a different data type.  
the next bit of code will build a small data frame that we can work with. Data frames have a lot of information that can be modified and it is a good  
idea to use R's built-in help to understand what you can do. to access it you do ?data.frame (replace data.frame with whatever you are curious about.)  

```{r data-frames}
x <- c(1,2,3)
y <- c("a","b","c")
z <- c(TRUE, FALSE, FALSE)
my_data <- data.frame(x,y,z,row.names=c("X","Y","Z"))
my_data
```

A very useful utility in R is the help feature. You can ask about any function and get some additional information by using it. Lets try it out with data.frames  
to get a better picture of what they really do. To access this you use ?<function_name>, try this now with data frame: ?data.frame and read through to find out  
how to get the names of the columns and ask what are the current column names for my_data.  

## Reading CSV files  
One of the more common actions you will do is to get data from a csv file, such as a survival curve for your experiments. R has several functions for reading in   tables from files, and it has two specific for csv files. read.csv and read.csv2. The biggest difference between those 2 functions is read.csv assumes the   seperator is a "," and that "." is used for decimal points and read.csv2 assumes ";" is the seperator and "," is the decimal point.  
In this repository there is a simple csv file that has some students, their current academic level and whether they like cake or pie. using the help  
function in R try and determine how to read this file into a data.frame (hint: check the examples from ?read.csv)  